---
title: "Assignment 2"
author: "Patrick"
date: "9/2/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install the Tidyverse

We begin by installing packages for easy data frame and date manipulation.
```{r installer, warning=FALSE, message=FALSE}
library("tidyverse")
library("lubridate")
```
##Loading and preprocessing the data

###1. Load the data (i.e. read.csv()) 

Reference the Github pages containing the activity.zip file.  The URL comes directly from 
[Dr. Peng's github repo](https://github.com/rdpeng/RepData_PeerAssessment1/blob/master/activity.zip) related to this assignment.

```{r loader}
datasource <-"https://github.com/rdpeng/RepData_PeerAssessment1/raw/master/activity.zip"
temp <- tempfile()
download.file(datasource, temp, mode = "wb")
unzip(temp, "activity.csv")
activity<-read.csv("activity.csv")
```

###2. Process/transform the data into a format suitable for your analysis

Use functions native to lubridate to cast activity date to a date from a factor.
```{r preprocess}
activity$date <- ymd(activity$date)
activity$weekday <- wday(activity$date, label = T)
activity$year <- year(activity$date)
activity$month <- month(activity$date)
activity$day <- day(activity$date)
activity$week<-week(activity$date)
activity$weekend<-as.factor(ifelse(activity$weekday %in% c("Sat", "Sun"), "weekend", "weekday"))
```

##What is mean total number of steps taken per day?

###1. Make a histogram of the total number of steps taken each day
```{r report}
hist(activity$steps,  breaks=20, col="gray", xlab="Time Interval", main="Frequency of Steps by Interval")
```

###2. Calculate and report the mean and median total number of steps taken per day

 We see in our summary below that the **mean** number of steps is **`r round(mean(activity$steps, na.rm = T))`** and the **median** is **`r median(activity$steps, na.rm = T)`**.

```{r summarize}
summary(activity)
```

##What is the average daily activity pattern?

###1. Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r timeseries}
activity %>%
  group_by(interval) %>%
  summarise(avgstep = mean(steps, na.rm = T)) %>%
  ggplot(., aes(x = interval, y = avgstep)) + geom_line(stat = "identity") +
  ggtitle("Average Steps by Interval")
```

###2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r maxinterval}
activity %>%
  group_by(interval) %>%
  summarise(avgstep = mean(steps, na.rm = T)) %>%
  arrange(desc(avgstep)) %>%
  head(., n = 1)
```

##Perform Mean Imputation on Missing Values
###1. Calculate and report the total number of missing values in the dataset
```{r narows}
sum(is.na(activity$steps))
```
###2. Create a new dataset using mean imputation
```{r imputate}
imputed.activity<-activity %>%
  group_by(interval) %>%
  mutate(steps = replace(steps, is.na(steps), mean(steps, na.rm=T)))
```
###3. Create a historgram and report the new mean/median of dataset with imputed values.
We observe there is no difference in the mean or median of the imputed dataset compared to the original.
```{r imputed summarize}
hist(imputed.activity$steps)
mean(imputed.activity$steps)
median(imputed.activity$steps)
summary(imputed.activity)
```

##Are there differences in activity patterns between weekdays and weekends?

###1. Visualize new weekdays variable.
```{r weekend facet}
imputed.activity %>%
  group_by(interval, weekend) %>%
  summarise(avgstep = mean(steps, na.rm = T)) %>%
  ggplot(., aes(x = interval, y = avgstep, group=weekend)) + geom_line(stat = "identity") +
  facet_grid(weekend ~.) +
  ggtitle("Average Steps by Interval by Weekday")
```
